---
title: "Residual error"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Residual error}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Residual error

PKPDsim can simulate residual errors in your observed data, which
can be done with the `res_var` argument to the `sim()` function.
This argument requires a `list()` with one or more of the
following components:

- `prop`: proportional error: $$y = \hat{y} \cdot (1 + \mathcal{N}(0, prop))$$
- `add`: additive error: $$y = \hat{y} + \mathcal{N}(0, add))$$
- `exp`: exponential error: $$y = \hat{y} \cdot e^{\mathcal{N}(0, exp)}$$

These list elements can be combined, e.g. for a combined proportional and additive
error model one would write: `res_var = list(prop = 0.1, add = 1)`, which
would give a 10% proportional error plus an additive error of 1 concentration unit.

Below are some examples of the `res_var` argument

```{r load-lib, echo=FALSE}
library(PKPDsim)
library(ggplot2)
```

Combined proportional and additive:

```{r example}
mod <- new_ode_model("pk_1cmt_iv")
reg <- new_regimen(
  amt = 1000,
  n = 5,
  interval = 12,
  type = "bolus"
)
sim1 <- sim(
  mod,
  parameters = list(CL = 5, V = 150),
  res_var = list(prop = 0.1, add = 1),
  regimen = reg,
  only_obs = TRUE
)
ggplot(sim1, aes(x = t, y = y)) +
  geom_point()
```

Exponential:

```{r exp}
sim2 <- sim(
  mod,
  parameters = list(CL = 5, V = 150),
  res_var = list(exp = 0.1),
  regimen = reg,
  only_obs = TRUE
)
```

Besides including the residual error at simulation time, there is also the option to
include it afterwards. For that, the function `add_ruv()` is useful.

```{r example-2}
sim3 <- sim1
sim3$y <- add_ruv(
  x = sim3$y, 
  ruv = list(
    prop = 0.1, 
    add = 1
  )
)
ggplot(sim3, aes(x = t, y = y)) +
  geom_point()
```
