---
title: "PKPD model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PKPD model}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
---

Below is an example of a simple indirect-response model. With PK-PD models, the inital state of the PD system often depends on specific model parts. We can define the state of the ODE system statically using the `A_init=` argument, but this will not take any parameters into account. However, we can also define initial states dynamically using the `state_init` argument, which allow you to specify the state of each compartment using code. 

```{r example-pkpd}
library("PKPDsim")
library("ggplot2")

p_pkpd <- list(
  CL = 5,
  V  = 50,
  KIN = .02,
  KOUT=.5,
  EFF = 0.2
)

r1 <- new_regimen(
  amt = 100,
  interval = 12,
  n = 5
)

pkpd <- new_ode_model(
  code = "
    dAdt[1] = -(CL/V) * A[1];
    conc = A[1]/V;
    dAdt[2] = KIN * 1/(1+EFF*conc) - KOUT*A[2];
  ",
  state_init = "A[2] = KIN/KOUT;"
)

dat <- sim(
  ode = pkpd,
  n_ind = 25,
  omega = cv_to_omega(
    par_cv = list(CL = 0.1, V = 0.1, KIN = .05, KOUT = 0.1),
    p_pkpd
  ),
  parameters = p_pkpd,
  regimen = r1,
  verbose = FALSE
)

ggplot(dat, aes(x = t, y = y, colour = factor(id))) +
  geom_line() +
  scale_colour_discrete(guide = "none") +
  facet_wrap(~comp, scales = "free")
```


# Combine PK and PD models

As shown above, a PK-PD model can be written as a single set of differential equations. However, we often develop PK and PD models separately and e.g. want to plug various PK models into existing PD models. In `PKPDsim` you can two or more model parts separately in a list to the `code` argument:

```{r separate-pk-pd}
pkpd <- new_ode_model(
  code = list(
    pk = "dAdt[1] = -(CL/V) * A[1]; conc = A[1]/V; ",
    pd = " dAdt[1] = KIN * 1/(1+EFF*conc) - KOUT*A[1]; "
  ),
  state_init = list(pd = "A[1] = KIN/KOUT;")
)
```

The above two systems of ODEs will then be combined into a single one.
