---
title: "Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Model

`new_ode_model` is the function that creates a new ODE model that can be used in the `sim()` command. It defines the ODE system and sets some attributes for the model. The model can be specified in three different ways:

- `model`: a string that references a model from the library included in `PKPDsim`. Examples in the current library are e.g. `pk_1cmt_oral`, `pk_2cmt_iv`. To show the available models, run `new_ode_model()` without any arguments.
- `code`: using code specyfing derivatives for ODE specified in *pseudo-R* code
- `file`: similar to `code`, but reads the code from a file

## Model from library

For example, a 1-compartment oral PK model can be obtained using:

```{r new-model}
library("PKPDsim")
pk1 <- new_ode_model(model = "pk_1cmt_oral")
```

Run the `new_ode_model()` function without arguments to see the currently available models:

```{r available-models, error=TRUE}
new_ode_model()
```

## Custom model from code

The custom model needs to be specified as a string or text block:

```{r custom-model}
pk1 <- new_ode_model(code = "
  dAdt[1] = -KA * A[1]
  dAdt[2] = +KA * A[1] -(CL/V) * A[2]
")
```

The input code should adhere to the follow rules:

- the derivaties in the ODE system are defined using `dAdt`
- array indices for the derivatives and compartments are indicated with `[ ]` and start at 1.
- equations are defined using `=`
- some expections apply, e.g. power functions (see below).

The input code is translated into a C++ function. You can check that the model compiled correctly by typing the model name on the R command line, which prints the model information:

```{r print-custom-model}
pk1
```

If you want even more detailed information, you can also print the actual C++ function that is used *under-the-hood* by specifying the `cpp_show_code=TRUE` argument to the `new_ode_model()` function.

## More custom model options

You can introduce new variables in your code, but you will have to define them using `declare_variables` argument too:

```{r custom-model-variables}
pk1 <- new_ode_model(code = "
  KEL = CL/V
  dAdt[1] = -KA * A[1]
  dAdt[2] = +KA * A[1] -KEL * A[2]
", declare_variables = c("KEL"))
```

Also, when you want to use covariates in your ODE system (more info on how to define covariates is in the *Covariates* vignette), you will have to define them, both in the code and in the function call:

```{r custom-model-covariates}
pk1 <- new_ode_model(code = "
  CLi = WT/70
  KEL = CLi/V
  dAdt[1] = -KA * A[1]
  dAdt[2] = +KA * A[1] -(CL*(WT/70)/V) * A[2]
", declare_variables = c("KEL", "CLi"), covariates = c("WT"))
```

One exception to the input code syntax is the definition of power functions. `PKPDsim` does not translate those from the *pseudo-R* code to valid C++ syntax automatically. C/C++ does not use the `^` to indicate power functions but uses the `pow(value, base)` function instead, so e.g. an allometric model should be written as:

```{r custom-model-power-functions}
pk1 <- new_ode_model(code = "
  CLi = CL * pow((WT/70), 0.75)
  dAdt[1] = -KA * A[1]
  dAdt[2] = +KA * A[1] -(CLi/V) * A[2]
", declare_variables = c("CLi"))
```

### Dosing / bioavailability

The default dosing compartment and bioavailability can be specified using the `dose` argument. By default, the dose will go into compartment `1`, with a bioavailability of `1`. The `bioav` element in the list can be either a number or a character string referring a parameter.

```{r bioav, eval=FALSE}
pk1 <- new_ode_model(code = "
  dAdt[1] = -KA * A[1]
  dAdt[2] = +KA * A[1] -(CL/V) * A[2]
",
dose = list(cmt = 1, bioav = "F1"))
```

For dosing based on mg/kg, at currently there is no solution for that using `new_regimen()`, although that might change in the future. The way to implement this at the moment is by scaling the dose by the "weight" covariate using the bioavailability:

```{r scale}
mod <- new_ode_model(code = "
  dAdt[1] = -(CL/V)*A[1];
",
dose = list(cmt = 1, bioav = "WT"),
obs = list(cmt = 1, scale = "V"),
covariates = list("WT" = new_covariate(value = 70)))
```

### Observations

The observation compartment can be set by specifying a list to the `obs` argument, with the elements `cmt` and `scale`.

```{r obs}
pk1 <- new_ode_model(code = "
  dAdt[1] = -KA * A[1]
  dAdt[2] = +KA * A[1] -(CL/V) * A[2]
", obs = list(cmt = 2, scale = "V"))
```

The `scale` can be either a parameter or a number, the `cmt` can only be a number.

*Note that the variables specified inside the differential equation block are not available as scaling parameters. E.g. for allometry you will have to redefine the scaled volume as follows:*

```{r scale-2}
pk1 <- new_ode_model(code = "
  Vi = V * (WT/70)
  dAdt[1] = -KA * A[1]
  dAdt[2] = +KA * A[1] -(CL/Vi) * A[2]
", obs = list(cmt = 2, scale = "V * (WT/70)"))
```

## Custom model from file

The same syntax as explained above applies, but using the `file=` argument the input code is read from the specified file. This is just a convenience function, i.e. it allows you to modularize your code, and separate models and R code.

```{r model-from-file, eval=FALSE}
pk1 <- new_ode_model(
  file = "pk_1cmt_oral_nonlin_v1.txt",
  declare_variables = c("KEL", "CLi")
)
```
