---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Installation

PKPDsim can be installed from CRAN:

```{r cran, eval=FALSE}
install.packages("PKPDsim")
```

Or you can install the development version from GitHub:

```{r gh, eval=FALSE}
devtools::install_github("InsightRX/PKPDsim")
```

# First simulation

The main simulation function in `PKPDsim` is `sim()`. To be able to simulate a dosing regimen for a specific drug, at least the following **three** arguments are required:

- `ode`: the compiled ODE model (created using the `new_ode_model()` function)
- `parameters`: a `list` of parameter values for the model
- `regimen`: the dosing regimen (created using the `new_regimen()` function)

The model library in PKPDsim contains a small library of built-in PK and PD models, but of course more interesting is its ability to handle user-specified ODE systems. However, as a first example, let's implement the most simple example from the library:

```{r load-lib, echo=FALSE}
library(PKPDsim)
library(ggplot2)
```

```{r intro-example}
p <- list(CL = 1, V  = 10, KA = 0.5)
pk1 <- new_ode_model("pk_1cmt_oral")
r1 <- new_regimen(
  amt = 100,
  n = 5,
  interval = 12
)
dat <- sim(
  ode = pk1,
  parameters = p,
  regimen = r1
)
```

You probably noticed that the `new_ode_model()`-step took a few seconds to finish, while the simulation itself was in the order of milliseconds. In `new_ode_model()`, the model is compiled from C++ source code to binary code, which takes a few seconds. However, this has to be done only once. After compilation, the model is then available to be used in `sim()` for as long as the R session is open.

Let's look at the output. `PKPDsim` will output data in the "long" format, i.e. one row per observed timepoint, and split by compartment:

```{r view-dat}
head(dat)
```

To check what output was produced by `sim()`, let's plot it (installation of `ggplot2` required).

```{r plot-example}
ggplot(dat, aes(x = t, y = y)) +
  geom_line() +
  facet_wrap(vars(comp))
```

# Custom ODE model

We can write the same exact model using custom ODE code. The following example will also only output the concentration (`only_obs = TRUE`), and not the amounts in each compartment. We're also going to use an infusion instead of bolus injection:

```{r custom}
pk2 <- new_ode_model(
  code = "dAdt[1] = -(CL/V) * A[1]",
  obs = list(cmt = 1, scale = "V"),
  dose = list(cmt = 1)
)
r2 <- new_regimen(
  amt = 100,
  n = 5,
  interval = 12,
  type = "infusion",
  t_inf = 2
)
dat2 <- sim(
  ode = pk2,
  parameters = p,
  regimen = r2,
  only_obs = TRUE
)
```

```{r plot-example2}
ggplot(dat2, aes(x = t, y = y)) +
  geom_line()
```

# Included models

PKPDsim includes definitions of several PK/PD models used in the InsightRX platform.

These models can be installed as R packages using the `model_from_api()` function.

```{r install-sample-models, eval=FALSE}
model_from_api(
  system.file("models", "pk_vanco_anderson.json5", package = "PKPDsim"),
  to_package = TRUE,
  install_all = TRUE
)
```

Please note that these models may differ from those used in production by
InsightRX. They are provided for demonstration purposes only, and should not be
used for dosing patients.

## Example usage

To simulate PK, provide the model and parameters to `sim()` along with
a dosing regimen and covariates.

```{r sample-model-example, eval=FALSE}
## Create dosing regimen and covariates
reg <- new_regimen(
  amt = 100,
  n = 3,
  interval = 12,
  type = "infusion",
  t_inf = 1
)

covs <- list(
  WT = new_covariate(4),
  PMA = new_covariate(42),
  CR = new_covariate(0.5),
  CL_HEMO = new_covariate(0)
)

## Perform simulation
sim(
  ode = pkvancoanderson::model(),
  parameters = pkvancoanderson::parameters(),
  regimen = reg,
  covariates = covs
)
```
