---
title: "Variability"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Variability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This can be implemented using the `n` and `omega` options. The `omega` (nomenclature borrowed from NONMEM) should be specified as either a **vector** defining the lower triangle of the BSV or $\Omega$ matrix, or as a **matrix** defining the full $\Omega$ matrix. An alternative option is to specify the between-subject variability as CV, using the `cv_to_omega()` function, but this assumes there is no correlation between individual parameters.

The following:

```{r load-lib, echo=FALSE}
library(PKPDsim)
```

```{r omega}
model <- new_ode_model("pk_1cmt_iv")
parameters <- list(CL = 5, V = 50)
regimen <- new_regimen(
  amt = 100,
  n = 3,
  interval = 12,
  type = "infusion",
  t_inf = 2
)

dat <- sim(
  ode = model,
  parameters = parameters,
  regimen = regimen,
  n = 50,
  omega = c(
    0.2, 
    0.05, 0.2
  )
)
```

will simulate out data for 50 patients, assuming an Omega matrix as defined above. Alternatively, the $\Omega$ could also have been defined as:

```{r omega-2}
dat <- sim(
  ode = model,
  parameters = parameters,
  regimen = regimen,
  n = 50,
  omega = matrix(c(
    0.2, 0.05, 
    0.05, 0.2
  ), ncol = 2)
)
```

Or also using the coefficient of variation and without correlation between parameters:

```{r omega-3}
dat <- sim(
  ode = model,
  parameters = parameters,
  regimen = regimen,
  n = 50,
  omega = cv_to_omega(list(
    CL = 0.1, 
    V = 0.1
  ))
)
```

Note that using the `cv_to_omega` function assumes the *CV is on the SD-scale* and not on the variance scale (and the definition of CV uses the assumption $1 + \eta \approx 1 * \exp(\eta)$).

# Variability distribution

By default, PKPDsim will assume exponential distribution of all parameters if `omega` is specified. If normal distribution is desired for all parameters, please use the `omega_type` argument:

```{r omega-normal}
dat <- sim(
  ode = model,
  parameters = parameters,
  regimen = regimen,
  n = 50,
  omega = c(
    0.2, 
    0.05, 0.1
  ),
  omega_type = c("normal", "normal")
)
```

# More flexible variability models

To allow more flexibility in how between-subject variability enters the model, there is an alternative way of specifying variability. This approach is very similar to the way variability is encoded in NONMEM, i.e. variability components (*eta's*) are added explicitly in the model code. In `PKPDsim` this means that *eta's* should be treated just like regular parameters, but with `0` mean and normal distribution. See example below for the simulation of bioavailability using the logit-distribution.

```{r bioav}
mod1 <- new_ode_model(
  code = "
    CLi = CL * exp(eta1)
    Vi = V * exp(eta2)
    F1i =  exp(F1 + eta3) / (1 + exp(F1 + eta3))
    dAdt[1] = -KA * A[1]
    dAdt[2] = KA * A[1] - (CLi/Vi) * A[2]
  ", 
  declare_variables = c("CLi", "Vi", "F1i"),
  obs = list(cmt = 2, scale = "V * exp(eta2)"),
  dose = list(cmt = 1, bioav = "F1i")
)
reg1 <- new_regimen(amt = 100, n = 2, interval = 12, type = "oral")
dat <- sim(
  ode = mod1,
  regimen  = reg1,
  parameters = list(
    eta1 = 0,
    eta2 = 0,
    eta3 = 0,
    CL = 5,
    V = 50,
    KA = .5,
    F1 = 0.8
  ),
  t_obs = c(0:48),
  omega = c(
    0.1, 
    0.05, 0.1, 
    0,    0,   0.1
  ),
  n = 100,
  omega_type = c("normal", "normal", "normal"),
  output_include = list("parameters" = TRUE, variables = TRUE),
  only_obs = TRUE
)
```

```{r plot}
library(ggplot2)
ggplot(dat, aes(x = t, y = y, group = id)) + geom_line()
ggplot(dat, aes(x = F1i)) + geom_histogram()
```
